\documentclass[
draft,
oneside,
% paper=B5,
% openright,
% DIV=13,
% BCOR=5mm,
]{scrbook}

\input{preamble}
\input{definitions}

\begin{document}
% \includepdf{frontpagethesis.pdf}
\frontmatter
\tableofcontents

\mainmatter
\chapter{Topological materials}
\include{sections/weylDirac}
\include{sections/typeii}

\chapter{Charge current from the conformal anomaly}
\include{sections/derivation_theory}
\include{sections/derivation_LL}
\include{sections/derivation_tiltless}
\include{sections/derivation_tilt}

\section{Discussion of results}
In the static and local limit $\lim_{\omega \to 0} \lim_{\vec{q}\to 0}$ the  transverse response function $\chi^{xy}$ of the charge current to a temperature perturbation
\begin{equation}
  J^x = \chi ^{xy} \frac{- \nabla^y T}{T}
\end{equation}
from a single Dirac point was found  to be
\begin{equation}
  \lim_{\omega \to 0} \lim_{\vec{q}\to 0}
  \chi ^{xy}
  =
  \gamma_{N}
  \frac{e^2 B v_F}{4 (2\pi )^2 \hbar },
\end{equation}
with $\gamma _N$ a prefactor dependent on how many landau levels are included in the final evaluation of the response function.
The response function is independent of the chirality $s$ of the Dirac point.
It was found that $\gamma _0 = 1, \gamma _{20} \approx 2$ and that the prefactor goes like $\log N$.

Firstly, the result differ slightly from that found by~\textcite{arjonaFingerprintsConformalAnomaly2019}
\begin{equation}
  \lim_{\omega \to 0} \lim_{\vec{q}\to 0}
  \chi ^{xy}
  =
  2 \gamma_{N}
  \frac{e^2 B v_F}{4 (2\pi )^2 \hbar },
\end{equation}
which differ by a factor of two.

Secondly, the sum will diverge as $N\to \infty $.
However, not all Landau levels are filled, and thus the sum should not be taken to all levels.
Similarly to a Quantum Hall effect, the number of filled bands, the filling factor $\nu $, is inverse proportional to the $B$-field strength
\begin{equation}
  \nu \propto \frac{1}{B}.
\end{equation}
Thus, we expect that the $N$-sum should be truncated at a Landau level, given by the filling factor $\nu $.
A detailed derivation of the exact truncation of the $N$-sum has not been done.
If a precise result for the numerical prefactor is found to be of importance, this should be straightforward.

The divergence is not discussed by~\citeauthor{arjonaFingerprintsConformalAnomaly2019}, where only the values of $N = 0$ and  $N=20$ are given, and the final result is that of $N=20$.
Furthermore, they state that the contributions from higher values of $N$ decrease very rapidly.
However, we found the contributions to go like $1 /x$, which is not decreasing rapidly enough to give a finite total contribution, thus giving the total contribution diverging logarithmically.
\todo{Say that we are communicating with them  to better understand their choice of truncation?}
Comparing our result with the different procedure done by~\citeauthor{chernodubGenerationNernstCurrent2018}~\cite{chernodubGenerationNernstCurrent2018}, the numerical prefactor found in our calculation including only the first term ($M=0$) coincides very well with the numerical prefactor found there, with a ratio of $16 / 18$.

\backmatter
\appendix
\include{sections/appendix-stressotherterm}
\include{sections/appendix-conformalsymmetry}

\printglossaries

\printbibliography

% \includepdf{endpagethesis.pdf}
\end{document}
